Tuning for Software Analytics: is it Really Necessary?

### 2015-08-18 tuningpredictor.tex
typo

\begin{document}
% \conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{ Analytics Without  Tuning Considered Harmful?}
% \numberofauthors{1}
\author{\alignauthor Wei Fu, Tim Menzies, Xipeng Shen \\
       \affaddr{Computer Science, North Carolina State University, Raleigh, USA}\\
       fuwei.ee, tim.menzies, xipengshen@gmail.com}
\maketitle 
\thispagestyle{plain}
\pagestyle{plain}
\begin{abstract}
One of the ``black arts'' of data mining is setting the tuning
parameters that control  the miner.  We offer a simple,
automatic, and very effective  method for finding those tunings.
For the purposes of learning
software defect predictors this  method can quickly
find  good tunings that  dramatically change   the performance of a learner.
For example,
in this paper we show   tunings that  alter detection  precision  
 from 2\% to 98\%.

These results prompt for a change to standard methods in software analytics.
At least for defect prediction, 
it is no longer enough to just run a data miner and present the result
{\em without}  conducting a tuning optimization study.
The implication for other kinds of  analytics is now  an open and pressing issue.

@@

In the $21^{st}$ century, it is now impossible
to manually browse all the available software project
data. The PROMISE repository of SE data has grown to 200+ projects~\cite{promise15}
and this is just one of over a dozen open-source repositories
that are readily available to researchers~\cite{rod12}.
For example, at the time of this writing (May  2015), our web searches show that Mozilla Firefox has over 1.1 million bug reports, and platforms such as GitHub host over 14 million projects.

Faced with this information overload,
researchers in empirical SE
use  data miners  to generate 
{\em defect predictors from static code measures}.
Such   measures can be
automatically extracted from the code base, with very little effort
even for very large software systems~\cite{nagappan05}.

@@

The results of this paper show that neither of these points is supportable, at least when learning
defect detectors from static code attributes since:
\be
\item
Tuning  defect predictors is {\em remarkably simple};
\item
And can {\em dramatically improve the performance}. 
\ee
More specifically,
this paper explores six research questions about tuning defect predictors for
static code measures:
\bi
\item RQ1: {\em Does   tuning    improve the performance scores of a predictor?} We will show below
 examples of truly dramatic improvement;
e.g.  precision changes from 2\% to 98\%.
\item RQ2: {\em Does tuning change conclusions on what learners are better than others?} 
Numerous recent SE papers (e.g. \cite{lessmann2008benchmarking,hall11,me07b}) claim that some learners are better than others. We show that at least
some of those conclusions are completely changed by tuning.
\item RQ3: {\em Does tuning change conclusions about what factors are most important in software engineering?} Numerous recent SE papers (e.g. ~\cite{bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,%
herzig2013predicting}) use data miners to conclude that {\em this}
is more important than {\em that} for (e.g.) reducing software project defects.
Given the  tuning results of this paper, we show that such conclusions need to be revisited.
\item  RQ4: {\em Is tuning easy?} We show that one of the simpler multi-objective optimizers
(differential evolution~\cite{storn1997differential}) works very well for tuning defect predictors. 
\item RQ5: {\em Is tuning impractically slow?}. We achieved dramatic improvements in the performance scores
of our data miners in less than 100 evaluations (!!); i.e., very
quickly.
\item RQ6: {\em Should data miners be used ``off-the-shelf'' with their default tunings?} 
At least for defect prediction from static code measures, our answer is an emphatic ``no'' (and
the implication for other kinds of software analytics is now an open and urgent question).
\ei
Before beginning, we digress for one  clarification.
This paper is {\em not} arguing that
software analytics is somehow wrong-headed or misguided.
In the age of the Internet and global access to software engineering data,
there exists the  problem of information overload. {\em Something} must be done to
allow analysts to make conclusions via an automatic analysis over a lot of data.
The results of this paper is that for a particular local context
(a specific data set and a specific goal) there exists  
methods for optimizing the conclusions reached in that context.  
Those conclusions
may not generalize to other contexts but this  is not a council for despair. 
As shown here, 
there  exists general methods for finding
local conclusions in a particular context. Further,
those
methods are  very simple to implement and very fast to execute.

\section{But is This a New Result?}

In  other fields, the impact of tuning is well understood. However, as we argue
in this section, this result is not properly  acknowledged  in the field of software analytics.

The formal definitions of data mining stresses that these learners are heuristic explorers
of a truly   vast space of the possible models within a data set~\cite{mitchell1982generalization}. For example, given 20 columns of boolean variables,
there are $2^{20}$ possible models. Data mining is practical since various heuristics
allow automatic algorithms to ignore most of those possibilities.

While that heuristic search makes data mining possible, it also means that tunings
can significantly affect what comes out of a data miner.
When we tune a data miner, what we are really doing is changing how a learner applies
its heuristics. This means tuned data miners use different heuristics, which means they ignore different possible models, which means they return different models.

The fact that {\em how} we learn changes {\em what} we learn is not acknowledged
in the field of software analytics.
There exists many research papers
that use data miners to   show that certain factors
are more influential than others for (say)
predicting defects. As shown below, such conclusions can be dramatically
changed by the tuning process since those  ``influential'' factors are very different pre- and post- tuning. Also, those factors tend to  change from project to project or if the goal
of the tuning is altered.
Hence, many old papers    need to be revisited  and perhaps revised~\cite{bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting}.  
For example, one of us (Menzies) used data miners
to assert that some factors were more important than others for predicting
successful software reuse~\cite{me02k}. That assertion should now be doubted since Menzies did not conduct a tuning study before reporting what factors the data miners
found were most influential.

Further, several  prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
that learnerX is better than learnerY for some software analytics task.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
software defect prediction~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst learner (CART) can perform just as well as the supposedly
best learner (Random Forest). Hence, all those prior results that ranked learners for software
analytics now need to be revisited (and perhaps revised).

Thirdly, it is a standard practice to use the default ``off-the-shelf'' tunings  for data mining tools (previously
we have defended that approach, arguing that it
encourages reproducibility~\cite{me15:book1}). That ``off-the-shelf''  policy
can no longer be condoned. For example, suppose our default
number of trees in  a Random Forest was   $F=100$ (which is the default in our implementation).
After tuning, we see that our optimizer often uses $F$ values that are nowhere near that default:

@@

{\scriptsize
\[F \in \left\{\begin{array}{l} 50, 58, 59, 63, 63, 66, 67, 73, 74, 74, \\
                                75, 77, 80, 82, 85, 96, 96, 97, 97, 101, \\
                                103, 107, 111, 111, 112, 120, 125, 125,  \\
                                130,130, 138, 144, 149, 150   \end{array}\right\}
\]}

@@

The  good news of this paper  is that {\em tuning is very easy}.
This result is both novel and unexpected.
A standard run of such evolutionary optimizers requires   thousands,
if not millions, of evaluations. However, in a result that we found startling, that  differential evolution can find useful settings for learners generating defect predictors
in less than 100 evaluations (i.e. very quickly).
Hence,   the ``problem'' (that
tuning changes the conclusions) is really
an exciting opportunity. At least for defect prediction,
 learners are very   amenable to tuning. Hence, 
 they are  also very
amenable to significant performance improvements. Given the low
number of evaluations required, then we assert that tuning
  should be standard practice
for anyone building defect predictors.

That said, the bad news is that, for defect
prediction, tuning dramatically changes
  what is learned from that data. Therefore, it is now an open and pressing research issue to check if
analytics without parameter tuning is considered {\em harmful} or, at the 
very least, {\em misleading}.
Clearly, we  must now revisit all prior results   based on
``off-the-shelf'' tunings.
Further,
it is no longer enough to just run a data miner and report the result
{\em without} first conducting a tuning optimization study.

@@

\section{Background Notes}

@@

There are many ways to build defect predictors
such as  CART~\cite{brieman00}, Random Forest~\cite{breiman84}, 
and WHERE~\cite{menzies2013local}.
For this study, we use the CART and Random Forest  from 
SciKitLearn~\cite{scikit-learn} (and
WHERE is available from
github.com/ai-se/where).
 We use  these algorithms, for the following reasons.
CART and Random Forest were mentioned in
a recent IEEE TSE paper by Lessmann et al.~\cite{lessmann2008benchmarking} that compared 21  
learners for  defect prediction.
That study ranked  CART  worst  and Random Forest as best.
In a demonstration of the impact of tuning,
this paper shows  we can {\em reverse} the conclusions of  Lessmann et al. such that CART
performs just as well as
 Random Forest.
This
 paper also presents results with WHERE-based learner since, as shown below,
it offers an interesting case study on the benefits of tuning.

@@

Our learners use the tuning parameters of \tab{parameters}. This section describes those parameters.

As shown in \tab{parameters}, default parameters for the Where-based learner are set using authors' engineering judgment and default parameters for CART and Random Forests are set by SciKitLearn. Those default parameters might not be the standard parameters for practical defect learners in reality, but it's a rule of thumb that most researchers would use just use the default parameters in the software without any tuning\cite{me07b,moser2008comparative,herzig2013predicting,zimmermann2007predicting}. The tuning range is set according to the corresponding default parameters. The reason for this is to justify whether tuning with a small range can get a performance improvement (\Menzies{is this correct?})

@@

These learners use different techniques to explore the splits:
\bi
\item
CART finds the attributes whose ranges contain rows with least variance in the number
of defects\footnote{If an attribute ranges $r_i$ is found in 
$n_i$ rows each with a  defect count variance of $v_i$, then CART seeks the attributes
whose ranges minimizes $\sum_i \left(\sqrt{v_i}\times n_i/(\sum_i n_i)\right)$.}.
\item
Random Forest    divides data like CART then  builds $F>1$  trees,
each time using some random subset of
the attributes. 
\item
When building the initial cluster tree, WHERE projects the data on to a dimension it synthesizes from the raw data using
a process analogous to principle component analysis\footnote{
PCA  synthesises  new
attributes $e_i, e_2,...$
that extends across the dimension of greatest  variance in the data  with attributes $d$.  
This process  combines
redundant  variables into a smaller set of variables  (so $e \ll d$) since those
redundancies become (approximately) parallel lines
in $e$ space. For all such redundancies \mbox{$i,j \in d$}, we 
can ignore $j$ 
since effects that change over $j$ also
change in the same way over $i$.
PCA is also useful for skipping over noisy variables from $d$-- these
variables are effectively ignored since    they  do not contribute to the variance in the data.}.
WHERE  divides  at the median point of that projection.
On recursion,
this generates the initial clustering tree, the leaves of which are clusters of  very similar examples. After that, when building 
the final decision tree, WHERE pretends its clusters are ``classes'', then 
asks the InfoGain of the
Fayyad-Irani discretizer~\cite{FayIra93Multi}, to rank the attriubutes, where {\em infoPrune} is used.
WHERE's final decision tree generator then ignores everything except the top   {\em infoPrune} percent of the sorted
attributes.
\ei
Some tuning parameters are learner specific:
\bi
\item
{\em Max\_feature} is used by
CART and Random Forest to select the number of attributes
used to build one tree.
CART's default is to use all the attributes while 
Random Forest usually selects the square root of the number
of attributes.
\item
  {\em Max\_leaf\_nodes} is the upper bound on leaf notes generated in a 
  Random Forest.
\item {\em Max\_depth} is the upper bound on the depth of the CART tree.  
 \item
WHERE's initial clustering
tree generation will always split up to {\em depthMin} number of branches.
After that, WHERE will only split data if the mean performance scores of the two halves
is ``trivially small'' (where ``trivially small'' is set by the   {\em wriggle} parameter). 
\item
WHERE's   {\em tree\_prune} setting controls how   
WHERE prunes back superflous parts of the final decision tree. 
If a decision sub-tree and its parent have the same 
majority cluster
(one that occurs most frequently), then if {\em tree\_prune} is enabled, we prune that decision sub-tree.
\ei

@@

 \subsubsection{Parametric Tuning Algorithms}
The  goal of this paper is to adjust the tuning parameters of \tab{parameters}
in order to   optimize (improve) some particular performance scores
generated by a particular learner being applied to  a particular data set.
For this task, we do not use traditional parametric numeric optimizer  
such as  gradient descent optimizers~\cite{saltelli00} that require models comprised of
differential functions (i.e. functions of real-valued variables whose derivative exists at each point in its domain).
This is impractical  for  our learners since their internal states are   not a smoothly differential continuous function.
Rather, learners being tuned  contain many regions with many different properties (tuning options can
drive the learner into very different modes with very different performance properties).

 \subsubsection{Non-Parametric Tuning Algorithms}

Non-parametric  optimizers   make no assumption
about the internal structure of the model.
There are many such optimizers:
\bi
\item
Simulated annealing~\cite{fea02a,me07f};
\item
A wide range of genetic algorithms~\cite{goldberg79} augmented by
techniques such as differential evolution~\cite{storn1997differential}, tabu search and scatter search~\cite{Glover1986563,Beausoleil2006426,Molina05sspmo:a,4455350};
\item
Particle swarm optimization~\cite{pan08}; 
\item 
Numerous ``decomposition approaches'' that uses
    heuristics to decompose the total space into   small problems,   then applies a
    simpler optimizer to each region including \mbox{$\varepsilon$-domination}~\cite{deb05} and MOEA/D ~\cite{zhang07};
    \item 
    Response surface methods~\cite{krall15,Zuluaga:13};
    \item
    And many others as well.
    \ei
From all the above methods, how do we select which optimizers to apply to tuning data miners?
Cohen~\cite{cohen95} advises comparing new 
 methods against the simplest possible alternative. 
Holte~\cite{holte93} recommends using very simple  learners
 as a
kind of ``scout'' for a  preliminary analysis of a data
set (to check if that data really requires a more
complex analysis).

To find our ``scout'',  we used engineering judgement to sort  the above algorithms from simplest to most complex.
Of the optimizers listed above, the simplest are simulated annealing (SA)  and 
differential evolution (DE), each of which can be coded in less than a page of some high-level scripting language. Our reading of the current literature is that there are more  advocates for
differential evolution than
  SA:
  \bi
  \item
  When the MOEA/D community requires a secondary optimizer, they often use  differential evolution~\cite{zhang07,5583335}.
  \item
 Vesterstrom and Thomsen~\cite{Vesterstrom04} found DE to be competitive with 
   particle swarm optimization and other GAs.
   \ei
DEs have been applied before for   parameter tuning (e.g. see~\cite{omran2005differential, chiha2012tuning}) but this is the first time they have been applied to
optimize defect prediction from static code attributes.  
The psuedocode for differential evolution is shown in Algorithm~\ref{alg:DE}.
In the following description, 
    superscript numbers denotes a line in that figure.

@@

   \caption{Data used in this experiment. 
   E.g., the top left data set has 20 defective classes out of 125 total.
   See \tion{dataa} for explanation of {\em training, tuning, testing} sets.
   }\label{fig:data1}
\end{table*}

>>>>

\begin{document}
% \conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{ Tuning for Software Analytics: is it Really Necessary?}
% \numberofauthors{1}
\author{\alignauthor Wei Fu, Tim Menzies, Xipeng Shen \\
       \affaddr{Computer Science, North Carolina State University, Raleigh, USA}\\
       fuwei.ee, tim.menzies, xipengshen@gmail.com}
\maketitle 
\thispagestyle{plain}
\pagestyle{plain}
\begin{abstract}
One of the ``black arts'' of data mining is setting the tuning
parameters that control  the miner.  We offer a simple,
automatic, and very effective  method for finding those tunings.

Contrary to our prior expectations, finding these tunings
 was remarkably simple: it only required tens, not thousands, of attempts
 to obtain very good results.
For example, when  learning
software defect predictors this  method can quickly
found tunings  that  alter detection  precision  
 from 2\% to 98\%.

Given that (1)~the improvements are so large, and (2)~the tuning is so simple,
these results prompt for a change to standard methods in software analytics.
At least for defect prediction, 
it is no longer enough to just run a data miner and present the result
{\em without}  conducting a tuning optimization study.
The implication for other kinds of  analytics is now  an open and pressing issue.

@@

In the $21^{st}$ century, it is  impossible
to manually browse all  available software project
data. The PROMISE repository of SE data has grown to 200+ projects~\cite{promise15}
and this is just one of over a dozen open-source repositories
that are readily available to researchers~\cite{rod12}.
For example, at the time of this writing (Aug  2015), our web searches show that Mozilla Firefox has over 1.1 million bug reports, and platforms such as GitHub host over 14 million projects.

Faced with this data overload,
researchers in empirical SE
use  data miners  to generate 
{\em defect predictors from static code measures}.
Such   measures can be
automatically extracted from the code base, with very little effort
even for very large software systems~\cite{nagappan05}.

@@

The results of this paper show that the above points are
false,  at least for
defect prediction from  code attributes:
\be
\item
Tuning  defect predictors is {\em remarkably simple};
\item
And can {\em dramatically improve the performance}. 
\ee
More specifically,
this paper explores six research questions about tuning defect predictors for
static code measures:
\bi
\item RQ1: {\em Does   tuning    improve the performance scores of a predictor?} We will show below
 examples of truly dramatic improvement:
 usually by 5 to 20\% and often by much more (in one extreme case, precision improved from 2\% to 98\%);
\item RQ2: {\em Does tuning change conclusions on what learners are better than others?} 
Recent SE papers~\cite{lessmann2008benchmarking,hall11} claim that some learners are better than others. 
Some of those conclusions are completely changed by tuning. 
\item RQ3: {\em Does tuning change conclusions about what factors are most important in software engineering?} Numerous recent SE papers (e.g. ~\cite{bell2013limited,rahman2013how,me02k,Moser:2008,zimmermann2007predicting,%
herzig2013predicting}) use data miners to conclude that {\em this}
is more important than {\em that} for reducing software project defects.
Given the  tuning results of this paper, we show that such conclusions need to be revisited.
\item  RQ4: {\em Is tuning easy?} We show that one of the simpler multi-objective optimizers
(differential evolution~\cite{storn1997differential}) works very well for tuning defect predictors. 
\item RQ5: {\em Is tuning impractically slow?} We achieved dramatic improvements in the performance scores
of our data miners in less than 100 evaluations (!); i.e., very
quickly.
\item RQ6: {\em Should data miners be used ``off-the-shelf'' with their default tunings?} 
For defect prediction from static code measures, our answer is an emphatic ``no'' (and
the implication for other kinds of  analytics is now an open and urgent question).
\ei
%Before beginning, we digress for one  clarification.
%This paper is {\em not} arguing that
%software analytics is somehow wrong-headed or misguided.
%In the age of the Internet and global access to software engineering data,
%there exists the  problem of information overload. {\em Something} must be %done to
%allow analysts to make conclusions via an automatic analysis over a lot of %data.
%The results of this paper is that for a particular local context
%(a specific data set and a specific goal) there exists  
%methods for optimizing the conclusions reached in that context.  
%Those conclusions
%may not generalize to other contexts but this  is not a council for despair. 
%As shown here, 
%there  exists general methods for finding
%local conclusions in a particular context. Further,
%those
%methods are  very simple to implement and very fast to execute.

\section{Background}

\subsection{Issues of Tuning: Important and Ignored?}
In  other fields, the impact of tuning is well understood~\cite{Bergstra2012}. 
Yet, as argued in this section, issues of tuning  are rarely or poorly addressed
in the defect prediction literature.

The formal definitions of data mining stresses that these learners are heuristic explorers
of a truly   vast space of the possible models within a data set~\cite{mitchell1982generalization}. For example, given 20 columns of boolean variables,
there are $2^{20}$ possible models. Data mining is practical since various heuristics
allow automatic algorithms to ignore most of those possibilities. 
While  heuristic search makes data mining possible, it also means that tunings
can significantly affect what comes out of a data miner.
When we tune a data miner, what we are really doing is changing how a learner applies
its heuristics. This means tuned data miners use different heuristics, which means they ignore different possible models, which means they return different models; i.e.  {\em how} we learn changes {\em what} we learn.

Are the impacts of tuning addressed in the defect prediction literature?
To answer that question,  in August 2015 we searched scholar.google.com for the conjunction of  "data mining" and "software engineering" and  "defect prediction".
After sorting by the citation count and discarding the non-SE papers (and those without a pdf link), we read over this sample
of  50 highly-cited SE defect prediction papers. 
What we found in that sample was that few authors
acknowledged the impact of tunings (exceptions: ~\cite{hall11,lessmann2008benchmarking}).
Overall,  80\% of papers in our sample {\em did not} adjust
the ``off-the-shelf'' configuration of the data miner (e.g. ~\cite{me07b,Moser:2008,herzig2013predicting,zimmermann2007predicting}). Of the remaining papers:
\bi
\item
Some papers in our sample  explored   data super-sampling~\cite{4271036} or data sub-sampling techniques via  automatic methods (e.g. ~\cite{Gao:2011,me07b,4271036,Kim:2011}) 
or via some domain principle (e.g. ~\cite{Moser:2008,Nagappan:2008,Hassan:2009}).
As an example of the latter, Nagappan et al.~\cite{Nagappan:2008} checked if metrics related to organizational structure were relatively more powerful for predicting software defects. 
However, it should be notes that  these studies varied the input data but
not the   ``off-the-shelf''   settings of the data miner.
\item
A few other papers did acknowledge that one data miner may not be appropriate
for all data sets.  Those papers tested  different  
``off-the-shelf'' data miners on the same data set.
For example, Elish et al.\cite{Elish2008649}  compared support vector
machines to other data miners for the purposes of defect prediction. SVM's execute via a ``kernel function'' which should be specially selected for different data sets and
the Elish et al. paper  makes no mention of any SVM tuning study.  
To be fair to Elish et al., we hasten to add that we
ourselves have  published
papers using ``off-the-shelf'' tunings~\cite{me07b} since,
prior to this paper it was unclear to us how to effectively
navigate the large space of possible tunings.
\ei
Over our entire sample, there was only  one paper that conducted a somewhat extensive tuning study.
Lessmann et al. tuned parameters for some of their algorithms using  a {\em grid search}; i.e. divide all $C$ configuration
options into $N$ values, then try all   $N^C$ combinations.
This is a slow approach-- we have explored grid search for 
defect prediction and found it takes days to terminate~\cite{me07b}.
Not only that, we found that grid search can miss
important optimizations~\cite{baker07}.
Every grid has ``gaps'' between each grid division which means
that a supposedly rigorous grid search can still miss
important configurations~\cite{Bergstra2012}. 
Bergstra and Bengio~\cite{Bergstra2012} comment for most data sets only a few of the tuning parameters really matter-- which means that
much of the runtimes associated with grid search is actually wasted.
Worse still, Bergstra and Bengio  comment that 
the 
important tunings are   different   for different
data sets-- a 
 phenomenon makes grid search a poor choice for configuring data mining
 algorithms for new data sets.

%We found only one paper in our sample from scholar.google.com 
%%that explored anything like the range of configuation values
%that we explore in this paper. Lessmann et %al.~\cite{lessmann2008benchmarking
%}
%explreod 22 learners
%Worse, in the rare paper
%that explores those options, it does so using a method that
%is known to perform poorly (the {\em grid search} method, discussed %below~\cite{Bergstra2012}).

%%Further, several  prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
%that learnerX is better than learnerY for some software analytics task.
%For example, a recent IEEE TSE article claimed that the 
%CART decision tree learner was far worse than Random Forests for
%software defect prediction~\cite{lessmann2008benchmarking}. 
%Such conclusions do not survive tuning.
%For example,
%after tuning, the worst learner (CART) can perform just as well as the supposedly
%best learner (Random Forest). Hence, all those prior results that ranked learners for software
%analytics now need to be revisited (and perhaps revised).

%Returning now to the issue
%of using  ``off-the-shelf'' tunings  for data mining tools: previously
%we have defended that approach, arguing that it
%encourages reproducibility~\cite{me15:book1}. Based on the results
%of this paper, it must be said that that ``off-the-shelf''  policy
%can no longer be condoned. For example, suppose our default
%number of trees in  a Random Forest was   $F=100$ (which is the default in our implementation).
%After tuning, we see that our optimizer often uses $F$ values that are nowhere near that default:

@@

@@

Since the Lessmann et al. paper, much progress has been made in 
configuration algorithms
and we can now report that  {\em finding useful tunings is very easy}.
This result is both novel and unexpected.
A standard run of grid search (and other  evolutionary algorithms)
is  that optimization requires   thousands,
if not millions, of evaluations.  However, in a result that we found startling, that  {\em differential evolution} (described below) can find useful settings for learners generating defect predictors
in less than 100 evaluations (i.e. very quickly).
Hence,   the ``problem'' (that
tuning changes the conclusions) is really
an exciting opportunity. At least for defect prediction,
 learners are very   amenable to tuning. Hence, 
 they are  also very
amenable to significant performance improvements. Given the low
number of evaluations required, then we assert that tuning
  should be standard practice
for anyone building defect predictors.

%That said, the bad news is that, for defect
%prediction, tuning dramatically changes
%  what is learned from that data. Therefore, it is now an open and %pressing research issue to check if
%analytics without parameter tuning is considered {\em harmful} or, at %the 
%very least, {\em misleading}.
%Clearly, we  must now revisit all prior results   based on
%``off-the-shelf'' tunings.
%Further,
%it is no longer enough to just run a data miner and report the result
%{\em without} first conducting a tuning optimization study.

@@

@@

There are many ways to build defect predictors
such as  CART~\cite{brieman00}, Random Forest~\cite{breiman84}, 
 WHERE~\cite{menzies2013local} and LR (logistic regression).
For this study, we use WHERE-based learner, CART, Random Forest and LR versions  from 
SciKitLearn~\cite{scikit-learn} (and
WHERE is available from
github.com/ai-se/where). 
 We use  these algorithms, for the following reasons.

CART and Random Forest were mentioned in
a recent IEEE TSE paper by Lessmann et al.~\cite{lessmann2008benchmarking} that compared 22  
learners for  defect prediction.
That study ranked  CART  worst  and Random Forest as best.
In a demonstration of the impact of tuning,
this paper shows  we can {\em reverse} the conclusions of  Lessmann et al. such that CART
performs just as well as
 Random Forest.

LR was  mentioned by Hall et al.~\cite{hall11}
as usually being as good or better as more complex learners (e.g.
Random Forests). In a finding that endorses the Hall et al. result,
we show that untuned LR performs better than 
untuned Random Forests (at least, for the data sets studied here). However,
we will show that tuning raises doubts about the optimality of the
Hall et al. recommendation.

Finally,  this
 paper uses WHERE since, as shown below,
it offers an interesting case study on the benefits of tuning.

%%%%%%%%%%%%%%%% list of parameters%%%%%%%%%%%%%%%%%%%%%
\renewcommand\arraystretch{1.2}
\begin{table*}[t!]
\scriptsize
  \centering
	\begin{tabular}{|c|c|c|c|l|}
	\cline{1-5}
	\begin{tabular}[c]{@{}c@{}}Learner \\ Name\end{tabular} & Parameters & Default &\begin{tabular}[c]{@{}c@{}}Tuning\\ Range\end{tabular}& 
\multicolumn{1}{c|}{Description} \\ \hline
	\multirow{8}{*}{\begin{tabular}[c]{@{}c@{}}Where-based\\ Learner\end{tabular}} 
	& threshold & 0.5 &[0.01,1]& The value to determine defective or not .\\ \cline{2-5} 
	& infoPrune & 0.33 &[0.01,1]& The percentage of features to consider for the best 
split to build its final decision tree. \\ \cline{2-5} 
	 & min\_sample\_split & 4& [1,10]& The minimum number of samples required to split an internal node of
its final  decision tree. \\ \cline{2-5} 
	 & min\_Size & 0.5 &[0.01,1]& \begin{tabular}[c]{@{}l@{}}Finds min\_samples\_leaf 
in the initial clustering tree  using  ${n\_samples}^ {min\_Size}$.
\end{tabular} \\ \cline{2-5} 
    & wriggle & 0.2 &[0.01, 1] & The threshold to determine which branch in  the initial clustering tree to be pruned\\ \cline{2-5}
	 & depthMin & 2 & [1,6]&The minimum depth of the initial clustering tree below which no pruning for the
clustering tree. \\ \cline{2-5} 
	 & depthMax & 10 &[1,20]& The maximum depth of the initial clustering tree. \\ \cline{2-5} 
	 & wherePrune & False &T/F& Whether or not to prune the initial clustering tree. \\ \cline{2-5}
	 & treePrune & True &T/F& Whether or not to prune the final decision tree. \\ \cline{2-5} 
\hline
\multirow{5}{*}{CART} & threshold & 0.5 &[0,1]& The value to determine defective or not. \\ \cline{2-5} 
	 & max\_feature & None &[0.01,1]& The number of features to consider when looking for the best 
split. \\ \cline{2-5}

	 & min\_sample\_split & 2 &[2,20]& The minimum number of samples required to split an 
internal node. \\ \cline{2-5} 
	 & min\_samples\_leaf & 1 & [1,20]&The minimum number of samples required to be at a leaf 
node. \\ \cline{2-5} 
     & max\_depth & None & [1, 50]& The maximum depth of the tree. \\
     \cline{1-5}  
       \multirow{5}{*}{\begin{tabular}[c]{@{}c@{}}Random \\ Forests\end{tabular}}  & threshold & 0.5 & [0.01,1] & The value to determine defective or not. \\ 
\cline{2-5} 
	 & max\_feature & None &[0.01,1]& The number of features to consider when looking for the best 
split. \\ \cline{2-5} 
	 & max\_leaf\_nodes & None &[1,50]& Grow trees with max\_leaf\_nodes in best-first fashion. \\ \cline{2-5} 
	 & min\_sample\_split & 2 &[2,20]& The minimum number of samples required to split an 
internal node. \\ \cline{2-5} 
	 & min\_samples\_leaf & 1 &[1,20]&The minimum number of samples required to be at a leaf 
node. \\ \cline{2-5} 
	 &  n\_estimators & 100 & [50,150]&The number of trees in the forest.\\ \cline{2-5}
	 \hline 
Logistic 	 Regression&\multicolumn{4}{l|}{This study uses untuned LR in order to check
a conclusion of~\cite{hall11}. }\\\hline

	\end{tabular}
    \caption {List of parameters tuned by this paper.}
\label{tab:parameters}
\end{table*}

@@

Our learners use the tuning parameters of \tab{parameters}. This section describes those parameters.
The default parameters for CART and Random Forests are set by 
the SciKitLearn authors and the
default parameters for WHERE-based learner are set via our own expert judgement.
When we say a learner is used ``off-the-shelf'', we mean
that they use the defaults shown in \tab{parameters}.

As to the value of those defaults, it could be argued that these defaults are 
not the best  parameters needed for practical defect prediction.
That said,  prior to this paper, two things were true:
\bi
\item 
Many data scientists in SE use the standard defaults
in their data miners, 
without   tuning (e.g.~\cite{me07b,Moser:2008,herzig2013predicting,zimmermann2007predicting}).
\item
The effort involved to adjust those tunings seemed so onerous, that
many researchers in this field were content to take our prior advice
of ``do not tune... it is just too hard''~\cite{me15:book1}.
\ei
As to the details of these learners, LR is a parametric
modeling approach. Given $f = \beta_0 + \sum_i\beta_ix_i$,
where $x_i$ is some measurement in a data set, and $\beta_i$
is learned via regression, LR
converts that into a function $0 \le g \le 1$
using $g=1/\left(1+e^{- f}\right)$. This function reports how much
we believe in a particular class.

@@

These learners use different techniques to explore the splits:
\bi
\item
CART finds the attributes whose ranges contain rows with least variance in the number
of defects\footnote{If an attribute ranges $r_i$ is found in 
$n_i$ rows each with a  defect count variance of $v_i$, then CART seeks the attributes
whose ranges minimizes $\sum_i \left(\sqrt{v_i}\times n_i/(\sum_i n_i)\right)$.}.
\item
Random Forest    divides data like CART then  builds $F>1$  trees,
each time using some random subset of
the attributes. 
\item
When building the initial cluster tree, WHERE projects the data on to a dimension it synthesizes from the raw data using
a process analogous to principle component analysis\footnote{
PCA  synthesises  new
attributes $e_i, e_2,...$
that extends across the dimension of greatest  variance in the data  with attributes $d$.  
This process  combines
redundant  variables into a smaller set of variables  (so $e \ll d$) since those
redundancies become (approximately) parallel lines
in $e$ space. For all such redundancies \mbox{$i,j \in d$}, we 
can ignore $j$ 
since effects that change over $j$ also
change in the same way over $i$.
PCA is also useful for skipping over noisy variables from $d$-- these
variables are effectively ignored since    they  do not contribute to the variance in the data.}.
WHERE  divides  at the median point of that projection.
On recursion,
this generates the initial clustering tree, the leaves of which are clusters of  very similar examples. After that, when building 
the final decision tree, WHERE pretends its clusters are ``classes'', then 
asks the InfoGain of the
Fayyad-Irani discretizer~\cite{FayIra93Multi}, to rank the attriubutes, where {\em infoPrune} is used.
WHERE's final decision tree generator then ignores everything except the top   {\em infoPrune} percent of the sorted
attributes.
\ei
Some tuning parameters are learner specific:
\bi
\item
{\em Max\_feature} is used by
CART and Random Forest to select the number of attributes
used to build one tree.
CART's default is to use all the attributes while 
Random Forest usually selects the square root of the number
of attributes.
\item
  {\em Max\_leaf\_nodes} is the upper bound on leaf notes generated in a 
  Random Forest.
\item {\em Max\_depth} is the upper bound on the depth of the CART tree.  
 \item
WHERE's  
tree generation will always split up to {\em depthMin} number of branches.
After that, WHERE will only split data if the mean performance scores of the two halves
is ``trivially small'' (where ``trivially small'' is set by the   {\em wriggle} parameter). 
\item
WHERE's   {\em tree\_prune} setting controls how   
WHERE prunes back superflous parts of the final decision tree. 
If a decision sub-tree and its parent have the same 
majority cluster
(one that occurs most frequently), then if {\em tree\_prune} is enabled, we prune that decision sub-tree.
\ei

@@

How should researchers select which optimizers to apply to tuning data miners?
Cohen~\cite{cohen95} advises comparing new 
 methods against the simplest possible alternative. 
Similarly, Holte~\cite{holte93} recommends using very simple  learners
 as a
kind of ``scout'' for a  preliminary analysis of a data
set (to check if that data really requires a more
complex analysis).
Accordingly,
to find our ``scout'',  we used engineering judgement to sort  candidate algorithms from simplest to  complex. For
example, here is a list of optimizers used widely in research:
{\em 
simulated annealing}~\cite{fea02a,me07f};
 various {\em genetic algorithms}~\cite{goldberg79} augmented by
techniques such as {\em differential evolution}~\cite{storn1997differential}, 
{\em tabu search} and {\em scatter search}~\cite{Glover1986563,Beausoleil2006426,Molina05sspmo:a,4455350};
{\em particle swarm optimization}~\cite{pan08}; 
numerous {\em decomposition} approaches that use
    heuristics to decompose the total space into   small problems,   then apply a
    {\em response surface methods}~\cite{krall15,Zuluaga:13}.
Of these,  the simplest are simulated annealing (SA)  and 
differential evolution (DE), each of which can be coded in less than a page of some high-level scripting language. Our reading of the current literature is that there are more  advocates for
differential evolution than
  SA. For example,  Vesterstrom and Thomsen~\cite{Vesterstrom04} found DE to be competitive with 
   particle swarm optimization and other GAs.

DEs have been applied before for   parameter tuning (e.g. see~\cite{omran2005differential, chiha2012tuning}) but this is the first time they have been applied to
optimize defect prediction from static code attributes.  
The psuedocode for differential evolution is shown in Algorithm~\ref{alg:DE}.
In the following description, 
    superscript numbers denotes a line in that figure.

@@

   \caption{Data used in this experiment. 
   E.g., the top left data set has 20 defective classes out of 125 total.
   See \tion{dataa} for explanation of {\em training, tuning, testing} sets.
   }\label{tab:data1}
\end{table*}
